Competing Constructions for German Measure Noun Phrases: from Usage Data to Experiment

### 2017-04-10 Latex/article.tex
Some Bildhauer comments.

This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
The study of alternations has a now longstanding tradition in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
This strand of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007}, and choices of variants are neither chosen deterministically nor fully at random.
Instead, multifactorial models are constructed, which incorporate influencing factors from diverse levels, including contextual ones.
The estimation of the model quantifies the influence the factors have on the probability that either variant is chosen.

For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the results with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
This position, which I adopt, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data \citep[486--487]{Dabrowska2016}.%
\footnote{I suggest that the word \textit{induce} would be more appropriate than \textit{deduce}.}
In fact, it would be surprising if we could!
Nobody assumes that we can induce mental representations from experiments, which -- as opposed to corpus studies -- even provide a direct access to the cognitive agent and offer much better options to control and experimental conditions and nuisance variables.
Instead, a theory of cognitive representation is pre-specified, and predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted.
Technical problems of statistical inference aside (see also Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
The simplest assumption would be that the same approach is adequate for corpus studies.

That said, the present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics, see \cite{DivjakArppe2013,Gries2003}.
It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternating variants.
The similarity of a given feature vector (in a concrete sentence where one of the variants is used) determines to a large extent which variant is chosen by speakers.
This entails that a variant of Prototype Theory with \textit{features} is assumed \citep{Rosch1978}, instead of the monolithic prototypes assumed in earlier versions.
While Prototype Theory is well suited for modeling constructional choices, it is just one similarity-based framework, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
Prototype Theory and Exemplar Theory differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or not (Exemplar Theory).
I am sceptical that corpus analysis alone could ever decide which theory is more adequate, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
\footnote{In \cite{DivjakArppe2013}, a fruitful attempt was made to show how both models form a converging picture based on corpus data.}
Prototype Theory (with feature abstractions) is simply preferred here because it fits the established alternation modeling paradigm, which relies on features being weighted (in a sense) in statistical models.
In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.

@@

In this section, I motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GLM(M)s -- does not require much motivation.
After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
All studies mentioned at the outset of Section~\ref{sec:cogocl} \citep{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010} use some form of regression\slash GLM(M).
Over the past few years, however, modified or alternative methods have been proposed.
From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016,Divjak2016a}, for example.
Conceptually, I see three points of discussion that should be kept apart.
First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data.
Third, and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.

As for the first point, this is not quite the place to discuss it fully.
The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
We can then calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
Clearly, in such quarrels between and among camps of philosophers of science and statisticians it is difficult for mere practitioners to take sides.

These quarrels relate to the second point, however.
\citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}). 
On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
\footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.}
Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora.
In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
I remain fully agnostic with respect to the question which approach is `right'.
The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.

@@

This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach, but in the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
Also, especially Neyman-style frequentism has well-known extensions to estimation (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).

@@

MLE, like inductive Bayesianism, conditions on the data in as much as it searches for the most likely set of parameters given the particular data.
What is more, Bayesian estimators are in fact based the Likelihood function and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.

@@

However, most similar studies (back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.

@@

I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind noun.
I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc. (such as \textit{glass} or \textit{bucket}).
In a similar vein, \citet[284]{Brems2003} considers nouns as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (\textit{sic!}) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).

In the case at hand, three different syntactic configurations need to be distinguished \wrt case assignment inside German measure noun phrases.
If the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
\footnote{If the kind noun is definite, the construction instantiates a true partitive.
Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \cite{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
}
Here, the kind noun is in the genitive, and I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).

@@

This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
\footnote{The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
While only an accusative is shown in (\ref{ex:intro:narrowapposition2}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.

@@

It was shown empirically in \cite{Hentschel1993} that such variants are de facto not acceptable.
Also, in my corpus sample, they simply did not occur.
Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using either corpus methods (Section~\ref{sec:corpusstudies}), and I consequently ignore them.

@@

Turning to (\ref{it:intro:plurl}), readers might have noticed that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (67 of 861 cases or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.%
\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.

As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear non-inflected.
They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which do not have plural forms at all.
%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
%A non-inflected variant \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected variant \textit{zwei Gläser roter Wein} co-exist.
%The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
In Section~\ref{sec:analyses}, degrees of grammaticalisation as a factor influencing the alternation will be discussed, however.

@@

This concludes the descriptive overview of the phenomenon.
I have shown that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
I turn to some more theory-oriented discussion in the next section.

@@

My analysis is based on the idea that the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} as instantiated in the \NACa\ and the \PGCa\ is ambiguous.
To show this, the first question is which noun constitutes the head of the whole construction.
This was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
However, the MNP-internal structure has more interesting cues to offer if the strong\slash weak inflection patterns of adjectives are taken into account.
In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.
%\footnote{The weak pattern has only two suffixes.
%The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
%The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves, however.
This is called the \textit{strong} inflectional pattern.
Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
On the one hand, they are lexical adjectives and function as attributive modifiers.
On the other hand, they are inflected like determiners, and they are the leftmost element in the NP, which is typical of determiners.
This unusual double nature of adjectives in NPs without determiners leads to a plausible probabilistic interpretation of the pattern shown in Table~\ref{tab:constructions}.
Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they have to use the \PGCa\ because, if there is a determiner, the PGC is the only option.
When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa.%
\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the GPC as a DP and in the NAC as an NP.}

>>>>

This paper deals with a morpho-syntactic alternation between two constructions that occurs only in a very specific type of measure noun phrase in German.
By \textit{alternation} I refer to a situation where two or more forms or constructions are available with no clear difference in acceptability, function, or meaning.
The study of alternations has a now longstanding tradition in cognitively oriented corpus linguistics (\citealp{BresnanEa2007,BresnanHay2010,BresnanFord2010,DivjakArppe2013,Gries2015,NessetJanda2010}, to name just a few publications).
This strand of research is based on the assumption that language is a probabilistic phenomenon \citep{Bresnan2007}, and choices of variants are neither chosen deterministically nor fully at random.
Instead, multifactorial models are constructed, which incorporate influencing factors from diverse levels, including contextual ones.
The estimation of the model coefficients quantifies the influence that the factors have on the probability that either variant is chosen.

For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the results with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} criticises this choice of words \textit{because it creates the impression that behavioral experimental data is inherently more valuable than textual data}, citing \cite{TummersEa2005}, who state that a corpus is \textit{a sample of spontaneous language use that is} (\textit{generally}) \textit{realized by native speakers}.
This position, which I adopt, does not, however, imply that we can in some way \textit{deduce mental representations from patterns of use}, \ie\ from corpus data, as  \citet[486--487]{Dabrowska2016} convincingly argues.%
\footnote{I suggest that the word \textit{induce} would be more appropriate than \textit{deduce}.}
It would be highly surprising if this were possible.
Nobody assumes that we can induce mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better options to control experimental conditions and nuisance variables.
Instead, a theory of cognitive representation is pre-specified, and predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted.
Technical problems of statistical inference aside (see also Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
The simplest assumption would be that the same approach is adequate for corpus studies.

That said, the present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics, see \cite{DivjakArppe2013,Gries2003}.
It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternating variants.
The similarity of a given feature vector (in a concrete sentence where one of the variants is used) determines to a large extent which variant is chosen by speakers.
This entails that a variant of Prototype Theory with \textit{features} is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
While Prototype Theory is well suited for modeling constructional choices, it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
Prototype Theory and Exemplar Theory differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or not (Exemplar Theory).
I am sceptical that corpus analysis alone could ever decide which theory is more adequate, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
\footnote{In \cite{DivjakArppe2013}, a fruitful attempt was made to show how both models form a converging picture based on corpus data.}
Prototype Theory (with feature abstractions) is simply preferred here because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.

@@

In this section, I motivate the choice of statistical models which I use later in Section~\ref{sec:corpusstudies}.
Readers might think that the method used here for modeling grammatical alternations -- namely (Hierarchical) Logistic Regression\slash Generalised Linear (Mixed) Models or GLM(M)s -- does not require much motivation.
After all, GLMMs have been established as the major tool in the analysis of alternation phenomena.
All studies mentioned at the outset of Section~\ref{sec:cogocl} use some form of regression\slash GLM(M).
Over the past few years, however, modified or alternative methods have been proposed.
From among these methods, I just make a few remarks on Bayesian estimation (see \citealp{GelmanEa2014}), as it was proposed in \cite{Levshina2016,Divjak2016a}, for example.
Conceptually, I see three points of discussion that should be kept apart.
First, Bayesian methods are sometimes touted as superior tools for scientific inference compared to frequentist methods.
Second, it has been proposed that the Bayesian interpretation of probability is more cognitively adequate for the modeling of linguistic data.
Third, and very specific to this paper, given established methods in the modeling of alternation and variation, it has to be decided whether so-called Bayesian methods lead to substantially different results.

As for the first point, this is not quite the place to discuss it fully.
The basic distinction is a philosophical one and related to the concepts of \textit{direct} and \textit{inverse probability} (\eg\ \citealp{Senn2011}).
Frequentists assume that models and parameters are fixed, for example a model specifying that a coin is fair.
We can then calculate for observed data (for example a measurement of $3$ heads in an experiment with $10$ tosses) how often such a result or a more extreme result would occur if the model were true and we repeated the experiment arbitrarily often.
This is essentially the frequentist notion of direct probability, \ie\ long-run frequencies under replication.
Standard tests in the Fisher and Neyman-Pearson traditions as well as Neyman confidence intervals are based on this concept of probability.
Bayesian approaches (in the now common interpretation), on the other hand, condition on the particular data and quantify inductively the probability of model parameters given the data.
The parameters are thus not fixed, and the probability is usually equated with researchers' posterior beliefs about model parameters.
There is actually a debate among Bayesians about the proper interpretation of Bayesian methods, and whether a notion of hypothesis testing is compatible (or even already contained) in the Bayesian approach.
In \citet[10]{GelmanShalizi2013}, the authors -- prominent Bayesians themselves -- acknowledge that a theory of statistical testing is a desideratum, and they state about the standard inductive interpretation of Bayesianism that \textit{most of this received view of Bayesian inference is wrong}, and they develop a Bayesian notion of $p$ values (see also \citealp{Mayo2013}, for a frequentist reply; also \citealp{Senn2011} on different strands of Bayesianism and their stance on inductive vs.\ deductive reasoning, and \citealp{Mayo2011}, for a critical reply to \citealp{Senn2011}).
Clearly, in such quarrels between and among camps of philosophers of science and statisticians, it is difficult for mere practitioners to take sides.

These quarrels relate to the second point, however.
\citet[301--302]{Divjak2016a} speaks favourably of Bayesian methods because the Bayesian concept of probability is more adequate for cognitive modeling compared to the frequentist one.
Her argument is part of a larger body of literature asking for cognitively plausible modeling techniques, for example Naive Discriminative Learning (NDL; \citealp{Baayen2011,BaayenEa2013,MilinEa2016,TheijssenEa2013}). 
On p.\ 303 of \cite{Divjak2016a}, the author goes on to explicitly mention NDL as well.
Yet, neither frequentist nor Bayesian methods were conceived as cognitive models, but as systems of inference for scientists (see above, and see also \citealp[302]{Divjak2016a}).
The fundamental question that lurks behind such arguments is how we interpret our statistical models (estimated on corpus data).
Are they inductive models of cognitive representations that also human learners would infer from being exposed to the corpus data?%
\footnote{In which case we are doing \textit{data science in language research} in the words of \citealp{MilinEa2016}.
I see this as standing in contradiction to the view advocated in \cite{Dabrowska2016} as cited above.}
Or are they tests of theories that are pre-specified and merely tested for predictive accuracy on linguistic output data contained in corpora?
In the former case, we adopt a strong \textit{corpus as input} hypothesis and should definitely resort to methods like NDL.
However, this would most likely require us to toss most previous work done in (cognitive) linguistics into the bin and to abandon all high-level generalisations that most existing studies have been based upon.
In fact, arguments to this extreme effect and against using high-level generalisation have been made, for example in \cite{BaayenEa2016}, \citet[299--300]{Divjak2016a}, \cite{RamscarPort2016}, \cite{TheijssenEa2013}.
In the latter and less extreme case, the cognitive commitment does, however, not necessarily extend to the statistical methods used.
These methods then do not need to be any more cognitively plausible than an ANOVA used to analyse the results from an experiment.
I view my own work (see Sections~\ref{sec:germanmeasurenps}--\ref{sec:experimental}) in the tradition of testing theories (which embrace high-level generalisations), and I agree to provisionally use Likelihood methods (see below).
I remain fully agnostic with respect to the question which approach is `right'.
The best strategy for cognitive linguistics as a field might be to cultivate many methods while making sure that each method is applied carefully and competently.

@@

This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach.
In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).

@@

MLE, like inductive Bayesianism, conditions on the particular data in as much as it searches for the most likely set of parameters given the data.
What is more, Bayesian estimators are in fact based the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.

@@

However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.

@@

I use the term \textit{measure noun phrase} (MNP) to refer a noun phrase (NP) in which a kind-denoting (count or mass) noun depends on another noun that specifies a quantity of the objects or the substance denoted by the kind noun.
I call the kind-denoting noun simply the \textit{kind noun} and the quantity-denoting noun the \textit{measure noun}.
For illustration purposes, in the English \textit{a cup of fine coffee}, \textit{cup} is the measure noun, and \textit{coffee} is the kind noun.
Measure nouns can be all sorts of nouns which genuinely denote a quantity (such as \textit{litre} or \textit{amount}) but also nouns denoting containers, collections, etc. (such as \textit{glass} or \textit{bucket}).
In a similar vein, \citet[284]{Brems2003} considers such nouns denoting containers etc.\ as measure nouns \textit{which, strictly speaking, do not designate a `measure', but display a more nebulous} (sic!) \textit{potential for quantification} (see also \citealp[530]{Koptjevskaja2001}, and \citealp[338]{Rutkowski2007}).

In the case at hand, three different syntactic configurations need to be distinguished \wrt case assignment inside German measure noun phrases.
If the kind noun forms an NP with a determiner, the construction resembles (and is usually called) a \textit{pseudo-partitive} (on partitives and pseudo-partitives see, e.\,g., \citealp{Barker1998,Selkirk1977,Stickney2007,Vos1999}; for a recent application of the terminology to German, see \citealp{Gerstenberger2015}).%
\footnote{If the kind noun is definite, the construction instantiates a true partitive.
Whereas partitives are constructions denoting a proper part-of relation as in \textit{a sip of the wine}, pseudo partitives -- albeit syntactically similar and diachronically related to partitives in many languages -- merely denote quantities and contain indefinite kind nouns as in \textit{a sip of wine}.
In the literature on German, some authors incorrectly call the pseudo-partitive a \textit{partitive} \citep{Hentschel1993} while some realise the difference and at least mention it \citep{Eschenbach1994,GallmannLindauer1994,Loebel1989,Zimmer2015}.
}
Here, the kind noun is in the genitive, and I refer to the construction in (\ref{ex:intro:pseudopartitive1}) as the \textit{Pseudo-partitive Genitive Construction} (PGC).

@@

This construction is usually classified as a \textit{Narrow Apposition Construction} \citep{Loebel1986}, henceforth NAC.%
\footnote{The construction as in (\ref{ex:intro:narrowapposition2}) is also referred to as the \textit{Direct Partitive Construction} for other Germanic languages in which the PGC with the synthetic genitive is not available.
This nomenclature makes sense in contrast to the \textit{Indirect Partitive Construction} with prepositional linkers translating to `of' -- \ie\ analytic genitives -- in such languages, see \cite{HankamerMikkelsen2008} for Danish.
For German, this terminology is not distinctive enough, which is why I use the terms NAC and PGC.}
Notice that the unavailability of the genitive on the kind noun can be seen as following from a rather quirky constraint that genitive NPs in German require the presence of some strongly case-marked element (determiner or adjective) in addition to the head noun in order to be acceptable (\citealp{GallmannLindauer1994,Schachtl1989}; see also \citealp[160]{Eisenberg2013b}).
%While only an accusative is shown in (\ref{ex:intro:narrowapposition2}), the kind noun obligatory agrees in case also with nominative and dative measure nouns.

@@

It was shown empirically in \cite{Hentschel1993} that such variants are de facto not acceptable.
Also, in my corpus sample, they simply did not occur.
Even if they are accepted by some speakers, their extremely low frequency makes it virtually impossible to study them using corpus linguistic methods (Section~\ref{sec:corpusstudies}), and I consequently ignore them.

@@

Turning to (\ref{it:intro:plurl}), readers might have noticed that so far only MNPs denoting quantities of substances (mass kind nouns such as \textit{ein Glas roter Wein\slash roten Weines} `a glass of red wine') have been discussed.
If the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns ($794$ of $861$ cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.%
\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.

As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear non-inflected.
They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
\footnote{Plurals \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs.
They cannot be used in MNPs.}
%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
%A non-inflected variant \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected variant \textit{zwei Gläser roter Wein} co-exist.
%The non-inflected variant denotes the quantity of wine fitting into two glasses, the inflected variant denotes two glasses filled with wine.}
I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
In Section~\ref{sec:analyses}, degrees of grammaticalisation as a factor influencing the alternation will be discussed, however.

@@

This concludes the descriptive overview of the phenomenon.
I have demonstrated that there is an alternation between two measure noun constructions in a narrow syntactic configuration (kind NP with an adjective but without a determiner), and that the two constructions differ in the case of the kind noun (case agreement with the measure noun or genitive).
I turn to some more theory-oriented discussion in the next section.

@@

My analysis is based on the idea that the relevant syntactic structure [N\Sub{1}~[A~N\Sub{2}]\Sub{NP\Sub{2}}]\Sub{NP\Sub{1}} as instantiated in the \NACa\ and the \PGCa\ is ambiguous.
To show this, the first question is which noun constitutes the head of the whole construction.
This was answered quite clearly by \cite{Loebel1986} already (see also \citealp[213]{Eschenbach1994}, and \citealp[16]{GallmannLindauer1994}).
Subject-verb agreement is always realised on the measure noun, and we can therefore assume that the measure noun is the head.
However, the MNP-internal structure has more interesting cues to offer if the strong\slash weak inflection patterns of adjectives are taken into account.
In NPs with a strongly inflected determiner, attributive adjectives inflect according to the massively syncretistic \textit{weak} pattern.
%\footnote{The weak pattern has only two suffixes.
%The structural cases (nominative and accusative) in the singular are marked with \textit{-e}, the oblique cases (dative and genitive) as well as the entire plural are marked with \textit{-en}.
%The only exception to this rule is the masculine accusative singular, which is also marked with \textit{-en}.}
If there is no determiner (as is the case in the alternating constructions), attributive adjectives inflect like determiners themselves, however.
This is called the \textit{strong} inflectional pattern.
Thus, the adjectives in the \NACa\ and the \PGCa\ have properties of adjectives as well as determiners.
On the one hand, they are lexical adjectives and function as attributive modifiers.
On the other hand, they are inflected like determiners, and they are the leftmost element in the NP, which is typical of determiners.
This unusual double nature of adjectives in NPs without determiners leads to a plausible probabilistic interpretation of the pattern shown in Table~\ref{tab:constructions}.
Whenever speakers classify the adjective in the kind noun phrase more as a determiner, they have to use the \PGCa\ because, if there is a determiner, the PGC is the only option.
When they classify the adjective more as an adjective, the kind NP has no determiner, and they have to use the \NACa.%
\footnote{While the generative analysis presented in \cite{Bhatt1990} cannot properly deal with probabilistic effects, Bhatt comes close to this interpretation by analysing the kind NP in the PGC as a DP and in the NAC as an NP.}
